---
title: "Rongkui_Chap13.1_HW"
author: "Rongkui Han"
date: "1/17/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rethinking)
data("reedfrogs")
d = reedfrogs
head(d)
```

#### 12M1   
Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size
treatment variables to the varying intercepts model. Consider models with either main effect alone,
both main effects, as well as a model including both and their interaction. Instead of focusing on
inferences about these two predictor variables, focus on the inferred variation across tanks. Explain
why it changes as it does across models.     

> Original model: 

```{r}
d$tank <- 1:nrow(d)

dat <- list(
  S = d$surv,
  N = d$density,
  tank = d$tank )

m13.2 <- ulam( 
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] ,
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 )
  ), data=dat , chains=4 , log_lik=TRUE )
```

> size model, main effect alone:    

```{r}
dat2 <- list(
  S = d$surv,
  N = d$density,
  tank = d$tank,
  Z = ifelse(d$size == 'small', 1, 2))

m13.3 <- ulam( 
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + b_z[Z], 
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ) ,
    b_z[Z] ~ dnorm(0, 1)
  ), data=dat2 , chains=4 , log_lik=TRUE )
```

> prediction model, main effect alone:    

```{r}
dat3 <- list(
  S = d$surv,
  N = d$density,
  tank = d$tank,
  pred = ifelse(d$pred == 'no', 1, 2))

m13.4 <- ulam( 
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] + b_p[pred], 
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ) ,
    b_p[pred] ~ dnorm(0, 1)
  ), data=dat3 , chains=4 , log_lik=TRUE )
```

> interaction model: 

```{r}
dat4 <- list(
  S = d$surv,
  N = d$density,
  tank = d$tank,
  Z = ifelse(d$size == 'small', 0, 1), 
  pred = ifelse(d$pred == 'no', 0, 1),
  interaction = ifelse(d$size == 'small', 0, 1)*ifelse(d$pred == 'no', 0, 1))

m13.5 <- ulam( 
  alist(
    S ~ dbinom( N , p ), 
    logit(p) <- a[tank] + b_z*Z + b_p*pred + b_int*interaction, 
    a[tank] ~ dnorm( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dexp( 1 ) ,
    b_z ~ dnorm(0, 1) ,
    b_p ~ dnorm(0,1) ,
    b_int ~ dnorm(0,1)
  ), data=dat4 , chains=4 , log_lik=TRUE )
```

#### 12M2    
Compare the models you fit just above, using WAIC. Can you reconcile the differences in
WAIC with the posterior distributions of the models?   
```{r}
compare(m13.2, m13.3, m13.4, m13.5)
```



#### 12M3   

Re-estimate the basic Reed frog varying intercept model, but now using a Cauchy distribution
in place of the Gaussian distribution for the varying intercepts. That is, fit this model:
si ~ Binomial(ni, pi)   
logit(pi) = atank[i]   
atank ~ Cauchy(a, s)   
a ~ Normal(0, 1)    
s ~ HalfCauchy(0, 1)    

Compare the posterior means of the intercepts, atank, to the posterior means produced in the chapter,
using the customary Gaussian prior. Can you explain the pattern of differences?     

```{r}
m12m3 <- ulam(
  alist(
    S ~ dbinom( N , p ) ,
    logit(p) <- a[tank] ,
    a[tank] ~ dcauchy( a_bar , sigma ) ,
    a_bar ~ dnorm( 0 , 1.5 ) ,
    sigma ~ dcauchy(0, 1)
  ), data=dat , chains=4 , log_lik=TRUE )
```

```{r}
precis(m12m3)
compare(m12m3, m13.2)
```

#### 12H2       

Return to the Trolley data, data(Trolley), from Chapter 12. Define and fit a varying intercepts
model for these data. Cluster intercepts on individual participants, as indicated by the unique values in the id variable. Include action, intention, and contact as ordinary terms. Compare
the varying intercepts model and a model that ignores individuals, using both WAIC and posterior
predictions. What is the impact of individual variation in these data?    

```{r}
data(Trolley)
d <- Trolley

dat <- list(
  R = d$response,
  A = d$action,
  I = d$intention,
  C = d$contact,
  id = as.numeric(d$id))

m12h2.1 <- ulam(
    alist(
      R ~ dordlogit( phi , cutpoints ),
      phi <- bA*A + bC*C + BI*I ,
      BI <- bI + bIA*A + bIC*C ,
      c(bA,bI,bC,bIA,bIC) ~ dnorm( 0 , 0.5 ),
      cutpoints ~ dnorm( 0 , 1.5 )
    ) , data=dat , chains=4 , cores=4, log_lik = TRUE )
    
```


```{r}
m12h2.2 <- ulam( alist(
  R ~ dordlogit( phi , cutpoints ),
  phi <- a[id] + bA*A + bC*C + BI*I ,
  BI <- bI + bIA*A + bIC*C , 
  a[id] ~ dnorm(0, sigma),  
  c(bA,bI,bC,bIA,bIC) ~ dnorm( 0 , 0.5 ),
  cutpoints ~ dnorm( 0 , 1.5 ),
  sigma ~ dexp(1)) ,
  data=dat, chains=4, log_lik = TRUE)
```


```{r}
compare(m12h2.1, m12h2.2)
```

```{r}

```

