---
title: "Rongkui_Chap14.7"
author: "Rongkui Han"
date: "7/6/2020"
output: html_document
---

You want to rest the hypothesis that the range size of carnivores is influenced by their body size (larger carnivores roam further).  You want to be sure that if you do see a relationship that is is not just due to shared evolutionary history.  That is you want to do a phylogenetic regression.    

1) Load the data as follows:    

```{r}
library(rethinking)
library(ape)
library(ade4)
library(dplyr)
library(tidyverse)
```

```{r}
data("carni70")
```

```{r}
tre = read.tree(text=carni70$tre)
d = carni70$tab
```

2) Note that the species name given in "rownames(d)" and in "tre$tip.label" do not match.  Fix that (e.g. using `str_replace()`)    
```{r}
tre$tip.label = str_replace(tre$tip.label, "\\.", "_")
```

3) Should either range or size be transformed?  
```{r}
ggplot(d, aes(x = size)) +
  geom_histogram(bins = 15)
ggplot(d, aes(x = range)) +
  geom_histogram(bins = 15)
```

> Will log transform size and scale transform range. 

```{r}
d['log_size'] = log(d$size)
d['scaled_range'] = range(d$range)
```

4) Does size predict range?  Evaluate this question without and with accounting for phylogenetic relationships.    

Without phylogenetic relationships:
```{r}
dat_list <- list( 
  N_spp = nrow(d),
  R = d$scaled_range,
  S = d$log_size,
  Imat = diag( nrow(d) )
)
mod_noPhy <- ulam(
  alist(
    R ~ multi_normal( mu , SIGMA ),
    mu <- a + bS*S ,
    matrix[N_spp,N_spp]: SIGMA <- Imat * sigma_sq,
    a ~ normal( 0 , 1 ),
    bS ~ normal( 0 , 0.5 ),
    sigma_sq ~ exponential( 1 )
  ), data=dat_list , chains=4 , cores=4)
precis( mod_noPhy )
```
```{r}
plot(precis(mod_noPhy))
```

> There is a small but significant positive correlation between size and range. 

With phylogenetic relatshipship, linear method:
```{r}
tree_trimmed <- keep.tip( tre, rownames(d) )
Dmat <- cophenetic( tree_trimmed )
spp_obs = rownames(d)
N_spp = nrow(d)

Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm)

dat_list$V <- V[ spp_obs , spp_obs ]
# convert to correlation matrix
dat_list$rel <- dat_list$V / max(V)

mod_Phy <- ulam(
  alist(
    R ~ multi_normal( mu , SIGMA ),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]: SIGMA <- rel * sigma_sq,
    a ~ normal( 0 , 1 ),
    bS ~ normal( 0 , 0.5 ),
    sigma_sq ~ exponential( 1 )
  ), data=dat_list , chains=4 , cores=4 )

precis( mod_Phy, depth = 3)

```

```{r}
plot(precis( mod_Phy, depth = 3))
```

> Size effect sill present. The estimated value of bS is similar to that of the phylogenetic-blind model. Sigma_sq got really large for some reason... 

With phylogenetic relationships, OU method:
```{r}
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)

mod_OU <- ulam(
  alist(
    R ~ multi_normal( mu , SIGMA ),
    mu <- a + bS*S ,
    matrix[N_spp,N_spp]: SIGMA <- cov_GPL1( Dmat , etasq , rhosq , 0.01 ),
    a ~ normal(0,1),
    bS ~ normal(0,0.5),
    etasq ~ half_normal(1,0.25),
    rhosq ~ half_normal(3,0.25)
  ), data=dat_list , chains=4 , cores=4)
precis( mod_OU, depth = 3)
```

```{r}
plot(precis( mod_OU ))
```


5) Discuss your findings.      

> Curiously, the phylogenetic OU model predicts stronger correlations between size and range. 
