---
title: "Rongkui_Chap12_HW"
author: "Rongkui Han"
date: "11/22/2019"
output: 
  html_document: 
    keep_md: yes
---

10H3
```{r}
library(rethinking)
library(MASS)
data("eagles")
head(eagles)

d = list(
  y = eagles$y,
  n = eagles$n,
  pr = ifelse(eagles$P == 'L', 1, 2),
  v = ifelse(eagles$V == 'L', 1, 2),
  a = ifelse(eagles$A == 'A', 1, 2)
)

m10.3 <- map2stan(
  alist(
  y ~ dbinom( n , p ),
  logit(p) <- alpha + b_pr[pr] + b_v[v] + b_a[a],
  alpha ~ dnorm(0, 10),
  b_pr[pr] ~ dnorm(0, 5),
  b_v[v] ~ dnorm(0, 5),
  b_a[a] ~ dnorm(0, 5)
) , data=d, chains=4, core = 4, iter = 4000)
plot(m10.3)
precis(m10.3 , depth=2)
```

```{r}
m10.3.1 <- map(
  alist(
  y ~ dbinom( n , p ),
  logit(p) <- alpha + b_pr[pr] + b_v[v] + b_a[a],
  alpha ~ dnorm(0, 10),
  b_pr[pr] ~ dnorm(0, 5),
  b_v[v] ~ dnorm(0, 5),
  b_a[a] ~ dnorm(0, 5)
) , data=d)
precis(m10.3.1 , depth=2 )
```

(b) Now interpret the estimates. If the quadratic approximation turned out okay, then it’s okay to
use the map estimates. Otherwise stick to map2stan estimates. Then plot the posterior predictions.
Compute and display both (1) the predicted probability of success and its 89% interval for each row (i)
in the data, as well as (2) the predicted success count and its 89% interval. What different information
does each type of posterior prediction provide?     
```{r}
d_ind = list(
  n = eagles$n,
  pr = ifelse(eagles$P == 'L', 1, 2),
  v = ifelse(eagles$V == 'L', 1, 2),
  a = ifelse(eagles$A == 'A', 1, 2)
)
post_pred = extract.samples(m10.3)
p = list()
par(mfrow = c(2,4))
for (i in 1:8) {
  p[[i]] = inv_logit(post_pred$alpha + post_pred$b_pr[,d_ind$pr[i]] + post_pred$b_v[,d_ind$v[i]] + post_pred$b_a[,d_ind$a[i]])
  hist(p[[i]])
}

y = list()
par(mfrow = c(2,4))
for (i in 1:8) {
  y[[i]] = rbinom(d_ind$n[i], prob = inv_logit(post_pred$alpha + post_pred$b_pr[,d_ind$pr[i]] + post_pred$b_v[,d_ind$v[i]] + post_pred$b_a[,d_ind$a[i]]), size = 10000)
  hist(y[[i]])
}
```

```{r}
post_pred2 = extract.samples(m10.3.1)
p2 = list()
par(mfrow = c(2,4))
for (i in 1:8) {
  p2[[i]] = inv_logit(post_pred2$alpha + post_pred2$b_pr[,d_ind$pr[i]] + post_pred2$b_v[,d_ind$v[i]] + post_pred2$b_a[,d_ind$a[i]])
  hist(p2[[i]])
}

y2 = list()
par(mfrow = c(2,4))
for (i in 1:8) {
  y2[[i]] = rbinom(d_ind$n[i], prob = inv_logit(post_pred2$alpha + post_pred2$b_pr[,d_ind$pr[i]] + post_pred2$b_v[,d_ind$v[i]] + post_pred2$b_a[,d_ind$a[i]]), size = 10000)
  hist(y2[[i]])
}
```

(c) Now try to improve the model. Consider an interaction between the pirate’s size and age (immature or adult). Compare this model to the previous one, using WAIC. Interpret.    
```{r}
d_c = list(
  y = eagles$y,
  n = eagles$n,
  pr = ifelse(eagles$P == 'L', 1, 2),
  v = ifelse(eagles$V == 'L', 1, 2),
  a = ifelse(eagles$A == 'A', 1, 2),
  inter = c(1, 1, 2, 2, 3, 3, 4, 4)
)

m10.3.c = map2stan(
  alist(
  y ~ dbinom( n , p ),
  logit(p) <- alpha + b_pr[pr] + b_v[v] + b_a[a] + b_inter[inter],
  alpha ~ dnorm(0, 10),
  b_pr[pr] ~ dnorm(0, 5),
  b_v[v] ~ dnorm(0, 5),
  b_a[a] ~ dnorm(0, 5),
  b_inter[inter] ~ dnorm(0, 5)
) , data=d_c, chains=4, core = 4, iter = 4000)
#plot(m10.3.c)
precis(m10.3.c , depth=2)
compare(m10.3, m10.3.c)
```

> Interaction seems to be important.  

11E3. When count data are zero-inflated, using a model that ignores zero-inflation will tend to induce which kind of inferential error?     

> Underestimation of key parameters? Overdispersion? Hiding effects of interest or producing spurious inferences?   

11E4. Over-dispersion is common in count data. Give an example of a natural process that might
produce over-dispersed counts. Can you also give an example of a process that might produce underdispersed
counts?    

> In the case of flower opening time (i.e. hour of day when flowers open), existing heterogeneity of duration of flower opening introduces over-dispersion of flower count.   

> I would imagine covariation between parameters can produce underdispersion of data. This could come from negative feedback loops common in biological processes. For example, plants that flower earlier in the day might have longer opening durationss, while later flowering plants havd shorter opening durations.     

11H6. The data in data(Fish) are records of visits to a national park. See ?Fish for details. The
question of interest is how many fish an average visitor takes per hour, when fishing. The problem is
that not everyone tried to fish, so the fish_caught numbers are zero-inflated. As with the monks
example in the chapter, there is a process that determines who is fishing (working) and another process
that determines fish per hour (manuscripts per day), conditional on fishing (working). We want
to model both. Otherwise we’ll end up with an underestimate of rate of fish extraction from the park.     

You will model these data using zero-inflated Poisson GLMs. Predict fish_caught as a function
of any of the other variables you think are relevant. One thing you must do, however, is use a proper
Poisson offset/exposure in the Poisson portion of the zero-inflated model. Then use the hours variable
to construct the offset. This will adjust the model for the differing amount of time individuals
spent in the park.    

```{r}
data(Fish)
head(Fish)

f = list(
  fish = Fish$fish_caught,
  l = ifelse(Fish$livebait == 1, 1, 2),
  camp = ifelse(Fish$camper ==1, 1, 2),
  per = Fish$persons,
  ch = Fish$child, 
  log_h = log(Fish$hours)
)
m11.6 <- ulam(
  alist(
    fish ~ dzipois( p , lambda ),
    logit(p) <- ap,
    log(lambda) <- log_h + b_l[l] + b_c[camp] + b_p*per + b_ch*ch,
    ap ~ dnorm( 0 , 1 ),
    b_l[l] ~ dnorm( 0 , 1 ),
    b_c[camp] ~ dnorm( 0 , 1 ),
    b_p ~ dnorm( 0 , 1 ),
    b_ch ~ dnorm( 0 , 1 )
  ) , data=f, chains = 4, core = 4, iter = 4000)
precis(m11.6)
```


11H1. In 2014, a paper was published that was entitled “Female hurricanes are deadlier than male
hurricanes.”180 As the title suggests, the paper claimed that hurricanes with female names have caused
greater loss of life, and the explanation given is that people unconsciously rate female hurricanes as
less dangerous and so are less likely to evacuate.
Statisticians severely criticized the paper after publication. Here, you’ll explore the complete data
used in the paper and consider the hypothesis that hurricanes with female names are deadlier. Load
the data with:


```{r}
data(Hurricanes)
head(Hurricanes)
?Hurricanes
```

In this problem, you’ll focus on predicting deaths using femininity of each hurricane’s name.
Fit and interpret the simplest possible model, a Poisson model of deaths using femininity as a
predictor. You can use map or map2stan. Compare the model to an intercept-only Poisson model of
deaths. How strong is the association between femininity of name and deaths? Which storms does
the model fit (retrodict) well? Which storms does it fit poorly?     

```{r}
h <- list( 
  d = Hurricanes$deaths,
  f = Hurricanes$femininity
)
m11.1 <- map2stan(
  alist(
    d ~ dpois( lambda ),
    lambda <- exp(f^b_f),
    b_f ~ dnorm(0,1)
), data=h , log_lik=TRUE )
precis(m11.1, 2)

m11.1.1 <- map2stan(
  alist(
    d ~ dpois( lambda ),
    lambda ~ dexp(1)
), data=h)
precis(m11.1.1, 2)

```

11H2. Counts are nearly always over-dispersed relative to Poisson. So fit a gamma-Poisson (aka
negative-binomial) model to predict deaths using femininity. Show that the over-dispersed model
no longer shows as precise a positive association between femininity and deaths, with an 89% interval
that overlaps zero. Can you explain why the association diminished in strength?      

```{r}

```



