---
title: "Rongkui_Chap14.1_HW"
author: "Rongkui Han"
date: "3/4/2020"
output: 
  html_document: 
    keep_md: yes
---

### 14E1
Add to the following model varying slopes on the predictor x.    
yi ~ Normal(mu_i, sigma)   
mu_i = alpha_group[i] + beta * xi   
alpha_group ~ Normal(alpha, sigma_alpha)    
alpha ~ Normal(0, 10)    
beta ~ Normal(0, 1)    
sigma ~ HalfCauchy(0, 2)     
sigma_alpha ~ HalfCauchy(0, 2)    

> New model:    
> yi ~ Normal(mu_i, sigma)   
> mu_i = alpha_group[i] + beta[i] * xi    
> [alpha_group[i], beta[i]] ~ MVNormal([alpha, beta], S)   
> S = (sigma_alpha, 0, 0, sigma_beta)_2x2 R (sigma_alpha, 0, 0, sigma_beta)_2x2    
> alpha ~ Normal(0, 10)    
> beta ~ Normal(0, 1)    
> sigma ~ HalfCauchy(0, 2)    
> sigma_alpha ~ HalfCauchy(0, 2)    
> sigma_beta ~ HalfCauchy(0, 2)    
> R ~ LKJcorr(2)    

The syntax for the multi-variate normal:     
```{r, eval = FALSE}
c(a_group, beta)[i] ~ multi_normal(c(a,b), Rho, sigma_i)
sigma_i ~ exponential(i)
Rho ~ lkj_corr(2)
```

### 14E2
Think up a context in which varying intercepts will be positively correlated with varying
slopes. Provide a mechanistic explanation for the correlation.    

> "The rich gets richer"... something like that I guess? The response variable would be amount of asset a person owns at different time points. 

### 14E3
When is it possible for a varying slopes model to have fewer effective parameters (as estimated
by WAIC or DIC) than the corresponding model with fixed (unpooled) slopes? Explain.    

> When the correlation is high...?    

### 14M1
Repeat the café robot simulation from the beginning of the chapter. This time, set rho to zero,
so that there is no correlation between intercepts and slopes. How does the posterior distribution of
the correlation reflect this change in the underlying simulation?     

```{r}
library(rethinking)
a <- 3.5 # average morning wait time 
b <- (-1) # average difference afternoon wait time
sigma_a <- 1 # std dev in intercepts
sigma_b <- 0.5 # std dev in slopes
rho <- 0 # correlation between intercepts and slopes
Mu <- c( a , b )
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2 )
sigmas <- c(sigma_a,sigma_b) # standard deviations 14.5
Rho <- matrix( c(1,rho,rho,1) , nrow=2 ) # correlation matrix
# now matrix multiply to get covariance matrix
Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)
N_cafes <- 20
library(MASS) 
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )
a_cafe <- vary_effects[,1] 
b_cafe <- vary_effects[,2]
plot( a_cafe , b_cafe , col='blue' , xlab="intercepts (a_cafe)" , ylab="slopes (b_cafe)" )
# overlay population distribution
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) ) {
  lines(ellipse(Sigma,centre=Mu,level=l),col=col.alpha("black",0.2))
}
```

```{r}
set.seed(22) 
N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )
mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5 # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )
```

```{r}
m14m1 <- ulam( 
  alist(
    wait ~ normal( mu , sigma ),
    mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
    c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
    a ~ normal(5,2),
    b ~ normal(-1,0.5),
    sigma_cafe ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ) , data=d , chains=4 , cores=4 )
```

```{r}
precis(m14m1, depth = 2)
plot(m14m1, depth = 2)
```


```{r}
post <- extract.samples(m14m1)
dens( post$Rho[,1,2] )
```

> posterior of rho is about 0. 

### 14M2
Fit this multilevel model to the simulated café data:
Wi ~ Normal(mu_i, sigma)   
mu_i = alpha_café[i] + beta_café[i] * Ai    
alpha_café ~ Normal(alpha, sigma_alpha)     
beta_café ~ Normal(beta, sigma_beta)    
alpha ~ Normal(0, 10)     
beta ~ Normal(0, 10)     
sigma_alpha ~ HalfCauchy(0, 1)    
sigma_beta ~ HalfCauchy(0, 1)    
sigma ~ HalfCauchy(0, 1)    
Use WAIC to compare this model to the model from the chapter, the one that uses a multi-variate
Gaussian prior. Explain the result.      
```{r}
m14.1 <- ulam( 
  alist(
    wait ~ normal( mu , sigma ),
    mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
    c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
    a ~ normal(5,2),
    b ~ normal(-1,0.5),
    sigma_cafe ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ) , data=d , chains=4 , cores=4, log_lik = TRUE)

m14m2 = ulam(
  alist(
    wait ~ normal( mu , sigma ),
    mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
    a_cafe[cafe] ~ dnorm(alpha, sigma_alpha),
    b_cafe[cafe] ~ dnorm(beta, sigma_beta),
    c(alpha, beta) ~ dnorm(0,10),
    c(sigma, sigma_alpha, sigma_beta) ~ dexp(1)
  ), data=d , chains=4 , cores=4, log_lik = TRUE)
```

```{r}
compare(m14.1, m14m2)
```

> They are quite similar when rho = 0. Re-doing simulation with a stronger correlation:   

```{r}
a <- 3.5 # average morning wait time 
b <- (-1) # average difference afternoon wait time
sigma_a <- 1 # std dev in intercepts
sigma_b <- 0.5 # std dev in slopes
rho <- 0.7 # correlation between intercepts and slopes
Mu <- c( a , b )
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2 )
sigmas <- c(sigma_a,sigma_b) # standard deviations 14.5
Rho <- matrix( c(1,rho,rho,1) , nrow=2 ) # correlation matrix
# now matrix multiply to get covariance matrix
Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)
N_cafes <- 20
library(MASS) 
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )
a_cafe <- vary_effects[,1] 
b_cafe <- vary_effects[,2]
plot( a_cafe , b_cafe , col='blue' , xlab="intercepts (a_cafe)" , ylab="slopes (b_cafe)" )
# overlay population distribution
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) ) {
  lines(ellipse(Sigma,centre=Mu,level=l),col=col.alpha("black",0.2))
}
set.seed(22) 
N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )
mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5 # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d_r0.7 <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )
```

```{r}
m14.1_r0.7 <- ulam( 
  alist(
    wait ~ normal( mu , sigma ),
    mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
    c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
    a ~ normal(5,2),
    b ~ normal(-1,0.5),
    sigma_cafe ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ) , data=d_r0.7 , chains=4 , cores=4, log_lik = TRUE)

m14m2_r0.7 = ulam(
  alist(
    wait ~ normal( mu , sigma ),
    mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
    a_cafe[cafe] ~ dnorm(alpha, sigma_alpha),
    b_cafe[cafe] ~ dnorm(beta, sigma_beta),
    c(alpha, beta) ~ dnorm(0,10),
    c(sigma, sigma_alpha, sigma_beta) ~ dexp(1)
  ), data=d_r0.7 , chains=4 , cores=4, log_lik = TRUE)
```

```{r}
compare(m14.1_r0.7, m14m2_r0.7)
```

> Hmmm still...? Did I do something wrong...?