---
title: "Rongkui_Chap16.1_HW"
author: "Rongkui Han"
date: "8/3/2020"
output: html_document
---

### 16M1. 

Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free
parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model.
How do they differ from those of m16.1?     

```{r}
library(rethinking) 
data(Howell1)
d <- Howell1
# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)
```

```{r, eval = FALSE}
m16.1 <- ulam( 
  alist(
    w ~ dlnorm( mu , sigma ),
    exp(mu) <- 3.141593 * k * p^2 * h^3,
    p ~ beta( 2 , 18 ),
    k ~ exponential( 0.5 ),
    sigma ~ exponential( 1 )
  ), data=d , chains=4 , cores=4 )

```

```{r, eval = FALSE}
m16.1_new <- ulam( 
  alist(
    w ~ dlnorm( mu , sigma ),
    exp(mu) <- 3.141593 * k * p^2 * h^exponent,
    p ~ beta( 2 , 18 ),
    k ~ exponential( 0.5 ),
    exponent ~ exponential( 0.5 ),
    sigma ~ exponential( 1 )
  ), data=d , chains=4 , cores=4 )
precis(m16.1_new)
```


###16M2. 

Conduct a prior predictive simulation for the cylinder height model. Begin with the priors
in the chapter. Do these produce reasonable prior height distributions? If not, which modifications
do you suggest?    
```{r}
N <- 1e4
p <- rbeta( N , 2 , 18 )
k <- rexp( N , 0.5 )

plot( NULL , xlim=c(0,1.5) , ylim=c(0,1) , xaxt="n" , xlab="h",ylab="w" )
at = c(0,0.25,0.5,0.75,1,1.25,1.5)
axis( 1 , at=at , labels=round(at*max(d$h), digit = 2) )
for ( i in 1:20 ) curve( (3.141593 * k[i] * p[i]^2 * x^3) , add=TRUE , col=grau() , lwd=1.5 )
```

```{r}

```

