---
title: "Rongkui_2.14_HW"
author: "Rongkui Han"
date: "2/13/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
data = read.csv("/Users/rongkui/Desktop/StatisticalRethinking/clarkia_transplant_data.csv")
head(data)
```


1. Fit a simple model with effects of temperature difference (temperature_diff_fall) on November germination (nov_germ). Temperature difference is already centered and scaled (i.e., negative values are the smallest temperature differences). Make sure to use the appropriate likelihood for the germination data (0 = no germ, 1  = germ).   

```{r}
library(rethinking)
d = list(
  tdf = data$temperature_diff_fall,
  ng = data$nov_germ
)

mod1 = ulam(
  alist(
    ng ~ dbinom(1, p),
    logit(p) <- u + a * tdf,
    u ~ dnorm(0, 1),
    a ~ dnorm(0, 1)
  ), data = d, chains = 4, cores = 4, log_lik = TRUE
)
precis(mod1, depth = 2)
```

2. Simulate from your priors to see if you've chosen reasonable priors, adjust them if necessary.

```{r}
set.seed(1)
prior = extract.prior(mod1)
plot( NULL , xlim=c(min(data$temperature_diff_fall),max(data$temperature_diff_fall)) , ylim=c(-0.2,1.2) ,
xlab="temp diff in fall" , ylab="germ prob", main = "Simulated prior" )
tdf_seq <- sample(data$temperature_diff_fall, 50)
mu <- link( mod1 , post=prior , data=data.frame(tdf = tdf_seq) )
dim(mu)
for ( i in 1:50 ) lines( tdf_seq , mu[i,] , col = col.alpha("black",0.3) )
```

These blocks were set up in the field, and had differences in soil depth, slope, and competitive environment. So maybe a model that includes block will describe the data better.

3. Fit a model that includes an effect of block (blk), with no pooling.    

```{r}
d = list(
  tdf = data$temperature_diff_fall,
  ng = data$nov_germ,
  blk = data$blk
)

mod3 = ulam(
  alist(ng ~ dbinom(1, p),
        logit(p) <- u + a * tdf + b[blk],
        u ~ dnorm(0,1),
        a  ~ dnorm(0,1), 
        b[blk] ~ dnorm(0, 1)), 
  data = d, chains = 4, cores = 4, log_lik = TRUE
)
precis(mod3, depth = 2)
plot(mod3, depth = 2)
compare(mod1, mod3)
```

Take out intersect:
```{r}
mod3b = ulam(
  alist(ng ~ dbinom(1, p),
        logit(p) <- a * tdf + b[blk],
        a  ~ dnorm(0,1), 
        b[blk] ~ dnorm(0, 1)), 
  data = d, chains = 4, cores = 4, log_lik = TRUE
)
precis(mod3b, depth = 2)
plot(mod3b, depth = 2)
compare(mod3, mod3b)
```


> So the block effects are real. 

4. Fit a model that includes block, and allows partial pooling.    

```{r}
mod4 = ulam(
  alist(
    ng ~ dbinom(1, p),
        logit(p) <- u + a * tdf + b[blk],
        u ~ dnorm(0,1),
        a  ~ dnorm(0,1), 
        b[blk] ~ dnorm(0, sigma_b), 
        sigma_b ~ dexp(1)),
  data = d, chains = 4, cores = 4, log_lik = TRUE
)
precis(mod4, depth = 2)
plot(mod4, depth = 2)
compare(mod4, mod3)
```


The experiment included many individuals from each of the 15 populations. So, each individual is not an independent representative of a given temperature, but might be similar to other plants from that population for reasons besides temperature.  

5. Build a model that accounts for this by including population (pop) and allowing partial pooling between populations A) without block, 

```{r}
d = list(
  tdf = data$temperature_diff_fall,
  ng = data$nov_germ,
  blk = data$blk,
  pop = data$pop
)

mod5a = ulam(
  alist(ng ~ dbinom(1, p),
        logit(p) <- u + a * tdf + b_p[pop],
        u ~ dnorm(0,1),
        a  ~ dnorm(0,1), 
        b_p[pop] ~ dnorm(0, sigma_p),
        sigma_p ~ dexp(1)), 
  data = d, chains = 4, cores = 4, log_lik = TRUE
)

precis(mod5a, depth = 2)
plot(mod5a, depth = 2)
```

and B) with block included as in the model above. How does including population affect the temperature estimate?   

```{r}
mod5b = ulam(
  alist(ng ~ dbinom(1, p),
        logit(p) <- u + a * tdf + b_p[pop] + b[blk],
        u ~ dnorm(0,1),
        a  ~ dnorm(0,1), 
        b_p[pop] ~ dnorm(0, sigma_p),
        b[blk] ~ dnorm(0, 1),
        sigma_p ~ dexp(1)), 
  data = d, chains = 4, cores = 4, log_lik = TRUE
)
precis(mod5b, depth = 2)
plot(mod5b, depth = 2)
```

6. Compare the five models you built using WAIC. Which fits best?

```{r}
compare(mod1, mod3, mod4, mod5a, mod5b)
```

> Model5b wins by a landslide.   

7. Plot effects of temperature difference for the average block,       
```{r}
post5b = extract.samples(mod5b)
p_link_bbar <- function( tdf ) {
  logodds <- with( post5b , u + as.vector(a * tdf) )
  return( inv_logit(logodds) )
}

seq = seq( min(data$temperature_diff_fall), max(data$temperature_diff_fall), length.out = 2000)

p_raw <- sapply(seq, function(i) p_link_bbar( i ))
p_mu <- apply(p_raw, 2, mean)
p_ci <- apply(p_raw, 2, PI)
plot( NULL , xlab="temperature difference in fall" , ylab="proportion germinated" , ylim=c(0,1) , xaxt="n" , xlim = c(min(data$temperature_diff_fall), max(data$temperature_diff_fall)) )
lines( seq , p_mu )
shade( p_ci , seq )
```

and also make a plot that includes the variability across blocks.    

> Oops the final model I made did not have a sigma_block (aka did not pool across block).   

```{r}
b_sim <- with( post5b , rnorm( length(post5b$u) , 0 , 1 ) )
p_link_bsim <- function( tdf ) {
  logodds <- with( post5b , u + as.vector(a * tdf) + b_sim)
  return( inv_logit(logodds) )
}

#seq = seq( min(data$temperature_diff_fall), max(data$temperature_diff_fall), length.out = 2000)

p_raw_sim <- sapply(seq, function(i) p_link_bsim( i ))
p_mu_sim <- apply(p_raw, 2, mean)
p_ci_sim <- apply(p_raw, 2, PI)
plot( NULL , xlab="temperature difference in fall" , ylab="proportion germinated" , ylim=c(0,1) , xaxt="n" , xlim = c(min(data$temperature_diff_fall), max(data$temperature_diff_fall)) )
lines( seq , p_mu_sim )
shade( p_ci_sim , seq )
```

