---
title: "Rongkui_Chap2_HW.Rms"
author: "Rongkui Han"
date: "4/18/2019"
output: 
  html_document: 
    keep_md: yes
---

### Easy.   

These problems use the samples from the posterior distribution for the globe tossing example. This code will give you a specific set of samples, so that you can check your answers exactly.    

R code 3.27    
```{r}
p_grid <- seq(from=0 , to=1 , length.out=1000) 
prior <- rep(1 , 1000)
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

Use the values in samples to answer the questions that follow.   
#### 3E1. How much posterior probability lies below p = 0.2?   
```{r}
library(ggplot2)
ggplot(data.frame(samples), aes(x = samples)) +
  geom_histogram()
p_lt_0.2 = sum(samples < 0.2)/length(samples)
```

> `r p_lt_0.2` of posterior probability lies below p = 0.2.    

#### 3E2. How much posterior probability lies above p = 0.8?   

```{r}
p_gt_0.8 = sum(samples > 0.8)/length(samples)
```

> `r p_gt_0.8` of posterior probability lies above p = 0.8.     

#### 3E3. How much posterior probability lies between p = 0.2 and p = 0.8?   

```{r}
p_gt_0.2_lt_0.8 = sum(samples > 0.2 & samples < 0.8)/length(samples)
```

> `r p_gt_0.2_lt_0.8` of posterior probability lies between p = 0.2 and p = 0.8.    

#### 3E4. 20% of the posterior probability lies below which value of p?   

```{r}
quant20 = quantile(samples, 0.2)
```

> 20% of the posterior probability lies below `r quant20`.   

#### 3E5. 20% of the posterior probability lies above which value of p?     

```{r}
quant80 = quantile(samples, 0.8)
```

> 20% of the posterior probability lies above `r quant80`.   

#### 3E6. Which values of p contain the narrowest interval equal to 66% of the posterior probability?     

```{r}
library(rethinking)
HPDI66 = HPDI(samples , prob=0.66)
```

> (`r HPDI66[1]`, `r HPDI66[2]`) contains the narrowest interval equal to 66% of the posterior probability.    

#### 3E7. Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?    

```{r}
PI66 = PI(samples , prob=0.66)
```

> (`r PI66[1]`, `r PI66[2]`) contains 66% of the posterior probability assuming equal posterior probability both below and above the interval.    


### Medium    
#### 3M1. Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.    

```{r}
likelihood.m1 <- dbinom(8 , size=15 , prob=p_grid)
posterior.m1 <- likelihood.m1 * prior
posterior.m1 <- posterior.m1 / sum(posterior.m1)
ggplot(data.frame(posterior = posterior.m1, p = p_grid), aes(x = p, y = posterior)) +
  geom_col()
```


#### 3M2. Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p.   

```{r}
samples.m2 = sample( p_grid , prob=posterior.m1 , size=1e4 , replace=TRUE)
ggplot(data.frame(samples.m2), aes(x = samples.m2)) +
  geom_histogram()
HPDI90 = HPDI(samples.m2 , prob=0.90)
```

> (`r HPDI90[1]`, `r HPDI90[2]`) is the 90% HPDI for p.    

#### 3M3. Construct a posterior predictive check for this model and data. This means simulate the distribution of samples, averaging over the posterior uncertainty in p. What is the probability of observing 8 water in 15 tosses?   

```{r}
w = rbinom(1e4 , size= 15, prob=samples.m2)
simplehist(w)
prob8 = sum(w == 8)/length(w)
```

> The probability of observing 8 water in 15 tosses is `r prob8`.  

#### 3M4. Using the posterior distribution constructed from the new (8/15) data, now calculate the probability of observing 6 water in 9 tosses.      

```{r}
w4 = rbinom(1e4 , size= 9, prob=samples.m2)
simplehist(w4)
prob6 = sum(w4 == 6)/length(w4)
```

> The probability of observing 6 water in 9 tosses is `r prob6`.  
