---
title: "Rongkui_Chap9_HW"
author: "Rongkui Han"
date: "10/2/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rethinking)
```


## Textbook exercises
### 9.3
```{r, 9.3}
# U: negative log probability
myU4 <- function(q, a = 0, b = 1, k = 0, d = 1) {
  muy <- q[1]
  mux <- q[2]
  U <- sum(dnorm(y, muy, 1, log = TRUE)) + sum(dnorm(x, mux, 1, log = TRUE)) + dnorm(muy, a, b, log = TRUE) + dnorm(mux, k, d, log = TRUE)
  return(-U)
}
```

```{r, 9.4}
# gradient function
# need vector of partial derivatives of U with respect to vector q
myU_grad4 <- function( q , a=0 , b=1 , k=0 , d=1 ) {
muy <- q[1]
mux <- q[2]
G1 <- sum( y - muy ) + (a - muy)/b^2 #dU/dmuy
G2 <- sum( x - mux ) + (k - mux)/d^2 #dU/dmuy
return( c( -G1 , -G2 ) ) # negative bc energy is neg-log-prob
}#
# test data
set.seed(7)
y <- rnorm(50)
x <- rnorm(50)
x <- as.numeric(scale(x))
y <- as.numeric(scale(y))
```


```{r, 9.5}
library(shape) # for fancy arrows
Q <- list()
Q$q <- c(-0.1,0.2)
pr <- 0.3
plot( NULL , ylab="muy" , xlab="mux" , xlim=c(-pr,pr) , ylim=c(-pr,pr) )
step <- 0.03
L <- 11 # 0.03/28 for U-turns --- 11 for working example
n_samples <- 4
path_col <- col.alpha("black",0.5)
points( Q$q[1] , Q$q[2] , pch=4 , col="black" )
for ( i in 1:n_samples ) {
  Q <- HMC2( myU4 , myU_grad4 , step , L , Q$q )
  if ( n_samples < 10 ) {
    for ( j in 1:L ) {
      K0 <- sum(Q$ptraj[j,]^2)/2 # kinetic energy
      lines( Q$traj[j:(j+1),1] , Q$traj[j:(j+1),2] , col=path_col , lwd=1+2*K0 )
    }
    points( Q$traj[1:L+1,] , pch=16 , col="white" , cex=0.35 )
    Arrows( Q$traj[L,1] , Q$traj[L,2] , Q$traj[L+1,1] , Q$traj[L+1,2] ,
    arr.length=0.35 , arr.adj = 0.7 )
    text( Q$traj[L+1,1] , Q$traj[L+1,2] , i , cex=0.8 , pos=4 , offset=0.4 )
  }
  points( Q$traj[L+1,1] , Q$traj[L+1,2] , pch=ifelse( Q$accept==1 , 16 , 1 ) ,
  col=ifelse( abs(Q$dH)>0.1 , "red" , "black" ) )
}
```

```{r, 9.6, eval = FALSE}
HMC2 <- function (U, grad_U, epsilon, L, current_q) { 
  q = current_q
  p = rnorm(length(q),0,1) # random flick - p is momentum.
  current_p = p
  # Make a half step for momentum at the beginning
  p = p - epsilon * grad_U(q) / 2
  # initialize bookkeeping - saves trajectory
  qtraj <- matrix(NA,nrow=L+1,ncol=length(q))
  ptraj <- qtraj
  qtraj[1,] <- current_q
  ptraj[1,] <- p
  # Alternate full steps for position and momentum 9.7
  for ( i in 1:L ) {
    q = q + epsilon * p # Full step for the position
    # Make a full step for the momentum, except at end of trajectory
    if ( i!=L ) {
      p = p - epsilon * grad_U(q)
      ptraj[i+1,] <- p
    }
  qtraj[i+1,] <- q
  }
  # Make a half step for momentum at the end
  p = p - epsilon * grad_U(q) / 2
  ptraj[L+1,] <- p
  # Negate momentum at end of trajectory to make the proposal symmetric
  p = -p
  # Evaluate potential and kinetic energies at start and end of trajectory
  current_U = U(current_q)
  current_K = sum(current_p^2) / 2
  proposed_U = U(q)
  proposed_K = sum(p^2) / 2
  # Accept or reject the state at end of trajectory, returning either
  # the position at the end of the trajectory or the initial position
  accept <- 0
  if (runif(1) < exp(current_U-proposed_U+current_K-proposed_K)) {
    new_q <- q # accept
    accept <- 1
  } else new_q <- current_q # reject
  return(list( q=new_q, traj=qtraj, ptraj=ptraj, accept=accept ))
}
```

### 9.4   
```{r, 9.9}
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[ complete.cases(d$rgdppc_2000) , ]
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )
head(dd)
```

```{r, 9.10}
m8.5 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dd )
precis( m8.5 , depth=2 )
```

```{r, 9.11}
dat_slim <- list(
  log_gdp_std = dd$log_gdp_std,
  rugged_std = dd$rugged_std,
  cid = as.integer( dd$cid )
)
str(dat_slim)
```

```{r, 9.12}
m9.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
) ,
  data=dat_slim , chains=1 )
```

```{r, 9.13}
precis( m9.1 , depth=2 )
```

```{r, 9.14}
m9.1 <- ulam( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dat_slim , chains=4 , cores=4 , iter=1000 )
show(m9.1)
```

```{r}
precis(m9.1, 2)
```

```{r}
pairs(m9.1)
```

```{r}
traceplot(m9.1, n_cols = 2)
trankplot(m9.1, n_cols = 2)
```

## Homework
### 8E3
> Discrete parameters. Discrete parameters don't form a surface the ball can glide one.   

### 8M1

Re-estimate the terrain ruggedness model from the chapter, but now using a uniform prior and
an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10)
and the exponential should be dexp(1). Do the different priors have any detectible influence on the
posterior distribution?    
```{r}
m8m1 <- ulam( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu ~ dunif( 0,10) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dat_slim , chains=4 , cores=4 , iter=1000 )
```


```{r}
precis(m8m1)
```

### 8M2   
The Cauchy and exponential priors from the terrain ruggedness model are very weak. They can be made more informative by reducing their scale. Compare the dcauchy and dexp priors for progressively smaller values of the scaling parameter. As these priors become stronger, how does each influence the posterior distribution?   

```{r}
m8m2.1 <- ulam( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu ~ dcauchy(0 , scale = 1) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dat_slim , chains=4 , cores=4 , iter=1000 , log_lik = TRUE )

m8m2.2 <- ulam( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu ~ dcauchy(0 , scale = 0.75) ,
    sigma ~ dexp( 0.75 )
  ) ,
  data=dat_slim , chains=4 , cores=4 , iter=1000 , log_lik = TRUE )

m8m2.3 <- ulam( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu ~ dcauchy(0 , scale = 0.5) ,
    sigma ~ dexp( 0.5 )
  ) ,
  data=dat_slim , chains=4 , cores=4 , iter=1000 , log_lik = TRUE )
```

```{r}
precis(m8m2.1)
precis(m8m2.2)
precis(m8m2.3)
WAIC(m8m2.1)
WAIC(m8m2.2)
WAIC(m8m2.3)
compare(m8m2.1, m8m2.2, m8m2.3)
```

### 8H1    
```{r}
mp <- ulam( 
  alist(
    a ~ dnorm(0,1),
    b ~ dcauchy(0,1)
  ),
  data=list(y=1),
  start=list(a=0,b=0),
  iter=1e4, warmup=100 )

precis(mp)
pairs(mp)
traceplot(mp)
```

> The model is sampling from the prior distributions of a and b.   
> A cauchy distribution has heavier tails than a normal distribution with the same mode, hence the extreme sample values.   

### 8H2    
```{r}
data(WaffleDivorce)
d <- WaffleDivorce
# standardize variables
d$A <- scale( d$MedianAgeMarriage )
d$D <- scale( d$Divorce )
d$M <- scale( d$Marriage ) 

m5.1 <- map2stan(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- a + bA * A ,
    a ~ dnorm( 0 , 0.2 ) ,
    bA ~ dnorm( 0 , 0.5 ) ,
    sigma ~ dexp( 1 )
  ) , data = d )


m5.2 <- map2stan(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- a + bM * M ,
    a ~ dnorm( 0 , 0.2 ) ,
    bM ~ dnorm( 0 , 0.5 ) ,
  sigma ~ dexp( 1 )
  ) , data = d )

m5.3 <- map2stan(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- a + bM*M + bA*A ,
    a ~ dnorm( 0 , 0.2 ) ,
    bM ~ dnorm( 0 , 0.5 ) ,
    bA ~ dnorm( 0 , 0.5 ) ,
    sigma ~ dexp( 1 )
  ) , data = d )

plot( coeftab(m5.1,m5.2,m5.3), par=c("bA","bM") )
compare(m5.1,m5.2,m5.3)
```

> Model 5.1 is better. Smaller WAIC, higher model weight.   

### Optional: 8H6   
```{r}
num_tosses <- 1e5
positions <- rep(0,num_tosses)
set.seed(1)
current <- sample(c('W','L'), 1)
for ( i in 1:num_tosses) {
  # record current position
  positions[i] <- current
  # flip coin to generate proposal
  proposal <- sample(c('W','L'), 1)
  # now make sure he loops around the archipelago
  if ( proposal == 'W' ) prob_move <- 0.7
  if ( proposal == 'L' ) prob_move <- 0.3
  # move?
  current <- ifelse( runif(1) < prob_move , proposal , current )
}
#positions
library(ggplot2)
pos = as.data.frame(positions)
ggplot(pos, aes(x = positions, fill = positions)) +
  geom_bar()
```



