---
title: "Roongkui_14.6_HW"
author: "Rongkui Han"
date: "6/28/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rethinking)
```

## 14M5

```{r}
data("Primates301")
data("Primates301_nex")
d <- Primates301 
d$name <- as.character(d$name)
dstan <- d[ complete.cases( d$group_size , d$body , d$brain ) , ]
spp_obs <- dstan$name
```

### No phylogenetic information
```{r}
dat_list <- list( 
  N_spp = nrow(dstan),
  M = standardize(log(dstan$body)),
  B = standardize(log(dstan$brain)),
  G = standardize(log(dstan$group_size)),
  Imat = diag( nrow(dstan) )
)
m14M5_1 <- ulam(
  alist(
    G ~ multi_normal( mu , SIGMA ),
    mu <- a + bM*M + bB*B,
    matrix[N_spp,N_spp]: SIGMA <- Imat * sigma_sq,
    a ~ normal( 0 , 1 ),
    c(bM,bB) ~ normal( 0 , 0.5 ),
    sigma_sq ~ exponential( 1 )
  ), data=dat_list , chains=4 , cores=4 )
precis( m14M5_1 )
```

```{r}
plot(precis( m14M5_1 ))
```

> Negatively correlated with brain size; positively correlated with body size. 

### With phylogenetic information

#### Calculate phylogentic distancee
```{r}
library(ape)
tree_trimmed <- keep.tip( Primates301_nex, spp_obs )
Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm)
Dmat <- cophenetic( tree_trimmed )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
```

> Linear

#### linear phylogenetic
```{r}
# put species in right order
dat_list$V <- V[ spp_obs , spp_obs ]
# convert to correlation matrix
dat_list$R <- dat_list$V / max(V)

m14M5_2 <- ulam(
  alist(
    G ~ multi_normal( mu , SIGMA ),
    mu <- a + bM*M + bB*B,
    matrix[N_spp,N_spp]: SIGMA <- R * sigma_sq,
    a ~ normal( 0 , 1 ),
    c(bM,bB) ~ normal( 0 , 0.5 ),
    sigma_sq ~ exponential( 1 )
  ), data=dat_list , chains=4 , cores=4 )

precis( m14M5_2)

```


#### OU method

```{r}

# Brownian motion model
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)

m14M5_2OU <- ulam(
  alist(
    G ~ multi_normal( mu , SIGMA ),
    mu <- a + bM*M + bB*B,
    matrix[N_spp,N_spp]: SIGMA <- cov_GPL1( Dmat , etasq , rhosq , 0.01 ),
    a ~ normal(0,1),
    c(bM,bB) ~ normal(0,0.5),
    etasq ~ half_normal(1,0.25),
    rhosq ~ half_normal(3,0.25)
  ), data=dat_list , chains=4 , cores=4)
precis( m14M5_2OU )
```




## 14H2
```{r}
data(Oxboys)
head(Oxboys)

d = list(
  height = Oxboys$height,
  age = Oxboys$age,
  subject = Oxboys$Subject)

m14H2 <- ulam( 
  alist(
    height ~ normal( mu , sigma ),
    mu <- a_sub[subject] + b_sub[subject]*age,
    c(a_sub,b_sub)[subject] ~ multi_normal( c(a,b) , Rho , sigma_subject),
    a ~ normal(149,5),
    b ~ normal(0,1),
    sigma_subject ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ) , data=d , chains=4 , cores=4 , log_lik = TRUE)
```

```{r}
precis(m14H2, depth = 3)
```

> How to see how much variaton is explained by each variable??

## 14H3:

There is positive correlation between intercepts (Rho[1,2]). Taller boys tend to also grow faster...?

