---
title: "Rongkui_Chap8_HW"
author: "Rongkui Han"
date: "7/31/2019"
output: 
  html_document: 
    keep_md: yes
---

### 7E1

(1) Temperature, sugar, salt, 
(2) major/race/gender
(3) functioning engine/tires/battery


### 7M3
In parts of North America, ravens depend upon wolves for their food. This is because ravens
are carnivorous but cannot usually kill or open carcasses of prey. Wolves however can and do kill
and tear open animals, and they tolerate ravens co-feeding at their kills. This species relationship
is generally described as a “species interaction.” Can you invent a hypothetical set of data on raven
population size in which this relationship would manifest as a statistical interaction? Do you think
the biological interaction could be linear? Why or why not?   

> probably not linear. Probably increase and then decrease.   

> raven.pop ~ a + b1 x prey.pop + b2 x wolf.pop x prey.pop

```{r}
prey.pop = rexp(1000, rate = 1)
wolf.pop = rnorm()
```

### 7H3   

Consider again the `data(rugged)` data on economic development and terrain ruggedness,
examined in this chapter. One of the African countries in that example, Seychelles, is far outside
the cloud of other nations, being a rare country with both relatively high GDP and high ruggedness.
Seychelles is also unusual, in that it is a group of islands far from the coast of mainland Africa, and
its main economic activity is tourism.   
One might suspect that this one nation is exerting a strong influence on the conclusions. In
this problem, I want you to drop Seychelles from the data and re-evaluate the hypothesis that the
relationship of African economies with ruggedness is different from that on other continents.   

(a) Begin by using map to fit just the interaction model:   

```{r}
library(rethinking)
library(tidyverse)
```

```{r}
data(rugged)
head(rugged)
d = rugged
d$log_gdp <- log( d$rgdppc_2000 )
# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ]
# rescale variables
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )
m1 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
data=dd )

```


```{r}
dd2 = dd[dd$country != 'Seychelles',]

m2 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
data=dd2 )
```


```{r}
precis( m1 , depth=2 )
```

```{r}
d.A1 <- dd[ dd$cont_africa==1 , ] # Africa
d.A0 <- dd[ dd$cont_africa==0 , ] # not Africa
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
# plot Africa - cid=1 8.17
plot( d.A1$rugged_std , d.A1$log_gdp_std , pch=16 , col=rangi2 ,
xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
xlim=c(0,1) )
mu <- link( m1 , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations")
# plot non-Africa - cid=2
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
xlim=c(0,1) )
mu <- link( m1 , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations")
```



```{r}
precis( m2 , depth=2 )
```

```{r}
d2.A1 <- dd2[ dd2$cont_africa==1 , ] # Africa
d2.A0 <- dd2[ dd2$cont_africa==0 , ] # not Africa
#rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
# plot Africa - cid=1 8.17
plot( d2.A1$rugged_std , d2.A1$log_gdp_std , pch=16 , col=rangi2 ,
xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
xlim=c(0,1) )
mu <- link( m2 , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations")
# plot non-Africa - cid=2
plot( d2.A0$rugged_std , d2.A0$log_gdp_std , pch=1 , col="black" ,
xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
xlim=c(0,1) )
mu <- link( m2 , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations")

```

```{r}
compare(m1, m2)
```


```{r}
head(dd)
```

