---
title: "HW_Jan31"
author: "Rongkui Han"
date: "1/31/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
data = read.csv("/Users/rongkui/Desktop/StatisticalRethinking/figure4phyE.csv")
head(data)
```

### Q1

#### a) subset the data for day 35     

```{r}
library(tidyverse)
data = data[data$day == 35,]
```

#### b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3    
```{r}
data["stem_length"] = (data$epi + data$int1 + data$int2 + data$int3)
```

#### c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.    
```{r}
data['flat2'] = ifelse(data$treatment == 'shade', data$flat + 6, data$flat)
data
mean(data$stem_length)
```

Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.    

### Q2
#### a) don't include flat.  Determine whether genotype, treatment, and their interaction are important predictors of stem_length     

```{r}
library(rethinking)

d = list(
  geno = as.numeric(data$genotype),
  trt = ifelse(data$treatment == "shade", 1, 0),
  sl = scale(data$stem_length)
)

mod2a = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt + inter[geno]*trt,
    g[geno]  ~ dnorm(0,1),
    t ~ dnorm(0,1),
    inter[geno] ~ dnorm(0,1),
    sigma_sl ~ dexp(1)
  ), data = d, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
precis(mod2a, depth = 2)
plot(mod2a, depth = 2)
```

```{r}
pairs(mod2a) #strong negative correlation between treatment effect and interaction effect
```

> Nothing seemed to have any effect...? Try without interaction   

```{r}
mod2a2 = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt,
    g[geno]  ~ dnorm(0, 1),
    t ~ dnorm(0,1),
    sigma_sl ~ dexp(1)
  ), data = d, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
precis(mod2a2, depth = 2)
plot(mod2a2, depth = 2)
```

```{r}
compare(mod2a, mod2a2)
```

> The one without interaction is much better. 

##### OR, you can keep the interaction and drop the main:    
```{r}
mod2a3 = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + inter[geno]*trt,
    g[geno]  ~ dnorm(0, 1),
    inter[geno] ~ dnorm(0,1),
    sigma_sl ~ dexp(1)
  ), data = d, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
precis(mod2a3, depth = 2)
plot(mod2a3, depth = 2)
```

```{r}
compare(mod2a, mod2a2, mod2a3)
```

> the main effect model is still better. But I guess it depends on what you are interested in. mod2a3 gives you a read out on response to shade in every genotype. 

#### b) starting with your best model from a), include flat without pooling     

```{r}
d2 = list(
  geno = as.numeric(data$genotype),
  trt = ifelse(data$treatment == "shade", 1, 0),
  flat = data$flat,
  sl = scale(data$stem_length)
)

mod2b = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt + bf[flat],
    g[geno]  ~ dnorm(0, 1),
    t ~ dnorm(0,1),
    bf[flat] ~dnorm(0,1),
    sigma_sl ~ dexp(1)
  ), data = d2, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
precis(mod2b, depth = 2)
plot(mod2b, depth = 2)
```

> There does seem to be a flat effect

```{r}
compare(mod2b, mod2a2)
```

> mod2b wins by a landslide. 

#### c) starting with your best model from a), use a hierarchical model that allows partial pooling across flats     

```{r}
mod2c = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt + bf[flat],
    g[geno]  ~ dnorm(0, 1),
    t ~ dnorm(0,1),
    bf[flat] ~ dnorm(flat_bar,sigma_flat),
    flat_bar ~ dnorm(0,1), 
    c(sigma_sl, sigma_flat) ~ dexp(1)
  ), data = d2, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
precis(mod2c, depth = 2)
plot(mod2c, depth = 2)
```

### Q3) Compare the models, which is preferred?

```{r}
compare(mod2b, mod2c)
```

> mod2b is slightly better. 

> We could center the pooleed effect at 0 and only look at sigma_fl

```{r}
mod2c2 = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt + bf[flat],
    g[geno]  ~ dnorm(0, 1),
    t ~ dnorm(0,1),
    bf[flat] ~ dnorm(0,sigma_flat),
    c(sigma_sl, sigma_flat) ~ dexp(1)
  ), data = d2, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
compare(mod2c, mod2c2)
```

> about the same

### Q4) Using the hierarchical model, make posterior predictions
#### a) for average cluster      

```{r}
post <- extract.samples(mod2c)
str(post)
```

```{r}
link_avg <- function(geno, trt) {
  with(post, g[,geno] + t*trt)
}
data35 <- data %>%
  filter(day==35) %>%
  mutate(stem_length=epi + int1 + int2 + int3,
         flat2=as.integer(as.factor(str_c(treatment, flat))),
         shade_i=ifelse(treatment=="sun", 0L, 1L),
         g_i= as.integer(factor(genotype,
                                   levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7")))) %>%
  select(genotype, treatment, g_i, shade_i, flat2, stem_length)
pred.df <- data35 %>% 
  select(-stem_length, -flat2) %>% 
  unique() %>%
  mutate(treatment=factor(treatment, levels = c("sun", "shade")) ) # so the plot order is correct
pred.df.avg <- pred.df %>%
  mutate(average_response=map2(g_i, shade_i, link_avg),
         pet_length=map_dbl(average_response, mean),
         low.89=map_dbl(average_response, ~ HPDI(.)[1]),
         high.89=map_dbl(average_response, ~ HPDI(.)[2]))
pred.df.avg
pred.df.avg %>%
  ggplot(aes(x=genotype, y=pet_length, ymin=low.89, ymax=high.89, fill=treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(width = .5, position = position_dodge(width=.9)) +
  ggtitle("prediction averaged across flat")
```

#### b) for same clusters   
```{r, eval = FALSE, include = FALSE}
#Shade condiction
## R code 13.36
p_link_abar_shade <- function( genotype , treatment_s=1, flat_factor=3) {
    stem_length_shade <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor] )
    return( stem_length_shade )
}
## R code 13.37
p_raw_shade <- sapply( 1:6 , function(i) p_link_abar_shade( i ) )
p_mu_shade <- apply( p_raw_shade , 2 , mean )
p_ci_shade <- apply( p_raw_shade , 2 , PI )
plot( NULL , main= "shade_3", xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_shade )
shade( p_ci_shade , 1:6 )
```


#### c) showing the "marginal" from cluster

```{r}
link_marg <- function(geno, trt) {
  with(post, g[,geno] + trt + rnorm(length(sigma_flat), 0, sigma_flat))
}
pred.df.marg <- pred.df %>%
  mutate(average_response=map2(g_i, shade_i, link_marg),
         pet_length=map_dbl(average_response, mean),
         low.89=map_dbl(average_response, ~ HPDI(.)[1]),
         high.89=map_dbl(average_response, ~ HPDI(.)[2]))
pred.df.marg
pred.df.marg %>%
  ggplot(aes(x=genotype, y=pet_length, ymin=low.89, ymax=high.89, fill=treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(width = .5, position = position_dodge(width=.9)) +
  ggtitle("prediction marginal to flat")
```


#### d) showing new clusters.

```{r}
plot.matrix <- sapply(pred.df.marg$average_response, function(x) x[1:50])
colnames(plot.matrix) <- str_c(pred.df.marg$genotype, "_", pred.df.marg$treatment)
plot.matrix <- t(plot.matrix)
plot.matrix[,1:10]
matplot(plot.matrix, type="l", xlab = colnames(plot.matrix))
```


### Q5) Reparameterize the model to help with divergent transitions (even if there aren't any)

```{r}
mod5 = ulam(
  alist(
    sl ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t*trt + bf[flat] + sigma_flat*norm, #could be g[geno] + t*trt + bf[flat]*sigma_flat, where bf[flat] ~ dnorm(0,1) and sigma_flat ~ dexp(1). But apparently Julin's didn't converge until he set sigma_flat~dnorm(0,1).
    g[geno]  ~ dnorm(0,1),
    t ~ dnorm(0,1),
    bf[flat] ~ dnorm(0,1),
    norm ~ dnorm(0,1), 
    c(sigma_sl, sigma_flat) ~ dexp(1)
  ), data = d2, chains = 4, cores = 4, iter = 2000, log_lik = TRUE
)
```

```{r}
plot(mod5, depth = 2)
```


```{r}
compare(mod5, mod2c)
```

> the reparameterized model is better.    

### Q6--optional
a) Which genotypes differ from MoneyMaker in Sun conditions?    
```{r}

```


b) Which genotypes differ from MoneyMaker in Shade conditions?
c) Which genotypes differ from MoneyMaker in their response to shade (difference in sun vs shade)?



