---
title: "HW_Jan31"
author: "Rongkui Han"
date: "1/31/2020"
output: html_document
---

```{r}
data = read.csv("/Users/rongkui/Desktop/StatisticalRethinking/figure4phyE.csv")
head(data)
```

### Q1

a) subset the data for day 35     

```{r}
data = data[data$day == 35,]
```

b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3    
```{r}
data["stem_length"] = (data$epi + data$int1 + data$int2 + data$int3)
```

c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.    
```{r}
data['flat2'] = ifelse(data$treatment == 'shade', data$flat + 6, data$flat)
data
```

Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.    

### Q2
a) don't include flat.  Determine whether genotype, treatment, and their interaction are important predictors of stem_length     

```{r}
library(rethinking)

d = list(
  geno = as.numeric(data$genotype),
  trt = as.numeric(data$treatment),
  interaction = as.numeric(as.factor(paste(data$genotype, data$treatment, sep = "_"))),
  sl = data$stem_length
)

mod2 = ulam(
  alist(
    stem_length ~ dnorm(mu, sigma_sl),
    mu <- g[geno] + t[trt] + int[interaction],
    g[geno]  ~ dnorm(0, 1),
    t[trt] ~ dnorm(0,1),
    int[interaction] ~ dnorm(0,1),
    sigma_sl ~ dexp(1)
  ), data = d, chains = 4, cores = 4, iter = 2000
)

precis(mod2)
```


b) starting with your best model from a), include flat without pooling
c) starting with your best model from a), use a hierarchical model that allows partial pooling across flats
```{r}

```

